---
title: 'TASK#6 : PREDICTION USING DECISION TREE ALGORITHM'
author: "SHAH MONA DEEPAK"
output:
  html_document:
    df_print: paged
---

## **TASK: TO CREATE A DECISION TREE CLASSIFIER AND VISUALIZE IT GRAPHICALLY** ##


## *WHAT IS A DECISION TREE?*
# Decision Tree is a Tree shaped Algorithm used to determine a course of action. Each branch of the tree represents a possible decision, occurrence or reaction.



# *STEP 1. LOAD THE DATASET.*

```{r}
## Here we are using IRIS Dataset for the given task.

data("iris")
data=iris
data
```

```{r}
dim(data)
```
```{r}
str(data)
```
**COMMENT: THE IRIS DATA HAS 150 OBSERVATIONS WITH 5 VARIABLES  'Sepal.Length' , 'Sepal.Width',  'Petal.Length' , 'Petal.Width' AND 'Species' WHICH IS A VECTOR OF CHARACTER TYPE- "SETOSA" ,"VERSICOLOR" ,"VIRGINICA" **  



# *STEP 2. DATA PARTITION.*

```{r}
set.seed(123)
sample=sample(2,nrow(data),replace=TRUE,prob=c(0.8,0.2)) #Prob=0.8 for training data, Prob=0.2 for test data.
table(sample)
```
```{r}
# We divide the data of 150 observations into 2 parts- Trained data consists of 121 observations while test data consists of 29 observations.
train=data[sample==1,]
test=data[sample==2,]
train

```
```{r}
dim(train)
```
```{r}
test
```
```{r}
dim(test)
```


# *STEP 3. DECISION TREE MODEL.*

```{r}
#Specify the libraries
library(party)
library(rpart)
library(rpart.plot)

#Develop the model
tree=rpart(Species~ Sepal.Length + Sepal.Width + Petal.Length + Petal.Width ,data=train)
tree
```


# *STEP 4. VISUALIZATION OF DECISION TREE.*

```{r}
rpart.plot(tree,type=5)
```

# *STEP 5. PREDICTION USING OBTAINED MODEL.*

```{r}
#Predict the observations from trained data into one of the Species.
predict(tree,train)
```
**COMMENT: ABOVE OUTPUT SUGGESTS THAT FOR A GIVEN OBSERVATION , THE TREE CLASSIFIES INTO ONE OF THE SPECIES WITH MAXIMUM PROBABILITY( SAY OBS. NO 144, THERE IS 0.9047 PROBABILITY THAT IT WILL BE CLASSIFIED INTO 'VIRGINICA' SPECIES.)**

```{r}
## We give observation at random to predict the class of Species it can belong.
New1=data.frame(Sepal.Length=c(4.6),Sepal.Width=c(3.1),Petal.Length=c(1.7),Petal.Width=c(0.2))
pred1=predict(tree,New1)
pred1
```
```{r}
New2=data.frame(Sepal.Length=c(5.0),Sepal.Width=c(2.8),Petal.Length=c(2.7),Petal.Width=c(0.8))
pred2=predict(tree,New2)
pred2
```

```{r}
New3=data.frame(Sepal.Length=c(5.0,5.2,3.6,4.3,2.7),Sepal.Width=c(2.8,4.3,1.7,2.3,1.4),Petal.Length=c(1.7,1.5,4.9,2.6,2.8),Petal.Width=c(0.8,0.2,0.5,0.7,0.9))
pred3=predict(tree,New3)
pred3
```

# *STEP 6. MISCLASSIFICATION ERRORS*

*MISCLASSIFICATION ERROR ON TRAINED DATA*
```{r}
p1=predict(tree,train,type="class")
t1=table(p1,train$Species)
t1
```
**COMMENT: THE OFF DIAGONAL ELEMENTS GIVES THE MISSCLASSICATION FIGURES FOR THE DATA IN TRAINING MODEL THAT HAS WRONGLY BEEN CLASSIFIED. IT SUGGESTS THAT 4 OBSERVATIONS HAVE BEEN CLASSIFIED TO VERSICOLOR WHILE ACTUALLY IT BELONGS TO VERGINICA SPECIES. **

```{r}
## TO CALCULATE MISSCLASSIFICATION (FOR TRAINED DATA):
m1=1-sum(diag(t1))/sum(t1)
m1
```
**COMMENT: THIS IMPLIES THAT MISCLASSIFICATION ERROR IS 4% WHILE THE MODEL IS 96% ACCURATE FOR THE TRAINING DATA.**




*MISCLASSIFICATION ERROR ON TEST DATA*
```{r}
p2=predict(tree,test,type="class")
t2=table(p2,test$Species)
t2
```
**COMMENT: THE OFF DIAGONAL ELEMENTS GIVES THE MISSCLASSICATION FIGURES FOR THE DATA IN TEST MODEL THAT HAS WRONGLY BEEN CLASSIFIED. IT SUGGESTS THAT 2 OBSERVATIONS HAVE BEEN CLASSIFIED TO VIRGINICA WHILE ACTUALLY IT BELONGS TO VERSICOLOR SPECIES. **


```{r}
## TO CALCULATE MISSCLASSIFICATION (FOR TEST DATA):
m2=1-sum(diag(t2))/sum(t2)
m2
```
**COMMENT: THIS IMPLIES THAT MISCLASSIFICATION ERROR IS 7% WHILE THE MODEL IS 93% ACCURATE FOR THE TEST DATA.**














